---
title: "Covid-Hosp-Be"
author: "Jean-Michel Bodart"
date: "09/04/2020"
output: html_document
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Covid-19 Hospitalizations in Belgium


```{r get_hosp_be, echo=FALSE, warning=FALSE}
suppressPackageStartupMessages({
library(dplyr)
library (tidyr)
library(httr)
library(lubridate)
library(ggplot2)
library(knitr)
#library(purrr)
})
reqdate <- format(seq.Date(as.Date(now()), by=-1, length.out=10), "%Y%m%d")
sources <- data.frame(reqdate, 
                      url = paste0("https://epistat.sciensano.be/Data/", 
                                   reqdate, 
                                   "/COVID19BE_HOSP_",
                                   reqdate,
                                   ".csv"),
                      stringsAsFactors = FALSE)
l <- lapply(sources$url, GET)
sources$status <- vapply(l, function(x){http_status(x)$reason}, "")
wh <- which(sources$status=="OK")
sources$lastmodif <- ''
if (length(wh)>0) {
  sources[wh,]$lastmodif <- vapply(l[wh], function(x){x$headers$`last-modified`}, "")
}
loc <- Sys.getlocale(category = "LC_TIME")
invisible(Sys.setlocale(category = "LC_TIME", locale = "C"))
sources$lastmodifdt <- as.POSIXct(NA)
sources$tz <- ''
sources$tz[wh] <- vapply(sources$lastmodif[wh], function(x){strsplit(x, " ", fixed = TRUE)[[1]][6]}, '')
src <- sources[wh,] %>% 
  mutate(lastmodifdt = as.POSIXct(lastmodif, tz=unique(tz), format="%a, %d %b %Y %H:%M:%S"))
src <- src %>%
  mutate(lastmodifdt = with_tz(lastmodifdt, tz="Europe/Brussels"))
# src <- src %>%
#   mutate(lastmodifdt = with_tz(lastmodifdt, tz="GMT"))
#cat(paste0(format(src$lastmodifdt, usetz = TRUE), "\n"))
sources$lastmodifdt[wh] <- src$lastmodifdt 
sources$tz <- NULL
#cat(paste0(format(src$lastmodifdt, usetz = TRUE), "\n"))
sources <- sources %>%
  mutate(lastmodifdt = with_tz(lastmodifdt, tz="GMT"))
#cat(paste0(format(src$lastmodifdt, usetz = TRUE), "\n"))

invisible(Sys.setlocale(category = "LC_TIME", locale = loc))

url <- sources$url[wh[1]]
dt <- sources$lastmodifdt[wh[1]]

```

#### Source: `r url`

Data last modified: `r format(dt, usetz = TRUE)`.


### COVID19BE_HOSP (HOSPITALISATIONS BY DATE AND PROVINCE) 

#### Dataset description: https://epistat.sciensano.be/COVID19BE_codebook.pdf.


Variable      | Definition 
------------- | ---------
DATE          | Date of data collection 
PROVINCE      | Province of reporting hospitals (including Brussels) 
REGION        | Region of reporting hospitals  
NR_REPORTING  | Number of hospitals reporting 
TOTAL_IN      | Total number of hospitalized patients at the moment of reporting (prevalence) 
TOTAL_IN_ICU  | Total number of hospitalized patients in ICU at the moment of reporting (prevalence) 
TOTAL_IN_RESP | Total number of hospitalized patients under respiratory support at the moment of reporting (prevalence) 
TOTAL_IN_ECMO | Total number of hospitalized patients on ECMO (Extracorporeal Membrane Oxygenation) at the moment of reporting (prevalence) 
NEW_IN        | Number of new hospital intakes in the last 24h (incidence) 
NEW_OUT       | Number of new hospital discharges (alive) in the last 24h (incidence) 


#### Newly Derived Variables

Variable | Definition 
-------------- | ---------
Full_reporting | Whether the number of hospitals reporting is equal to the maximal number of hospitals reporting 
TOTAL_IN_CHG   | (Total number of hospitalized patients at the moment of reporting) - (Total number of hospitalized patients the previous day)
NEW_DELTA      | (Number of new hospital intakes) - (new hospital discharges (alive) in the last 24h) 


```{r hosp_be, echo=FALSE, warning=FALSE}
# hosp <- data.table::fread("https://epistat.sciensano.be/Data/20200409/COVID19BE_HOSP_20200409.csv")

hosp0 <- data.table::fread(sources$url[wh[1]]) %>%
  mutate(DATE = as.Date(DATE), NEW_DELTA = NEW_IN - NEW_OUT) %>%
  arrange(REGION, PROVINCE, DATE)

# hosp <- hosp0 %>%
#   left_join(
#     select(hosp0, REGION, PROVINCE, DATE, starts_with("TOTAL")) %>%
#       mutate(DATE = DATE + 1) %>%
#       rename_all(
#         list(~ stringr::str_replace_all(., 'TOTAL_', 'PRV_'))
#       ),
#     by = c("REGION", "PROVINCE", "DATE")
#     )  

hosp <- hosp0 %>% 
  group_by(REGION, PROVINCE) %>%
  arrange(DATE, .by_group = TRUE) %>%
  mutate_at(
        vars(starts_with("TOTAL_")), # for all variables starting with "TOTAL_"
        funs(CHG = . - lag(.))       # compute change from previous record
                                     # and store as variable with added suffix "_CHG"
      ) %>%
  ungroup()
     

hospt <- pivot_longer(data = hosp
                    ,cols = -REGION & -PROVINCE & -DATE & - NR_REPORTING
                    ,names_to = "category"
                    ,values_to = "count") %>%
  mutate(country = "Belgium") %>%
  arrange(country, REGION, PROVINCE, DATE, category) %>%
  select(country, REGION, PROVINCE, DATE, NR_REPORTING, category, count)

province <- hospt %>%
  group_by(country, REGION, PROVINCE, DATE, category) %>%
  summarise(NR_REPORTING = sum(NR_REPORTING), count = sum(count)) %>%
  ungroup() %>%
  select(country, REGION, PROVINCE, DATE, NR_REPORTING, category, count)

region <- hospt %>%
  group_by(country, REGION, DATE, category) %>%
  summarise(NR_REPORTING = sum(NR_REPORTING), count = sum(count)) %>%
  mutate(PROVINCE="All") %>%
  ungroup() %>%
  select(country, REGION, PROVINCE, DATE, NR_REPORTING, category, count) %>%
  rbind(province)

country0 <- hospt %>%
  group_by(country, DATE, category) %>%
  summarise(NR_REPORTING = sum(NR_REPORTING), count = sum(count)) %>%
  mutate(REGION="All", PROVINCE="All") %>%
  ungroup() %>%
  select(country, REGION, PROVINCE, DATE, NR_REPORTING, category, count) %>%
  rbind(region) %>%
  arrange(country, REGION, PROVINCE, DATE, NR_REPORTING, category)

country <- country0  %>%
  group_by(country, REGION, PROVINCE) %>%
  summarize(MAX_REPORTING = max(NR_REPORTING)) %>%
  full_join(country0) %>%
  mutate(full_reporting = (MAX_REPORTING == NR_REPORTING)) %>%
  rename_all(stringr::str_to_title)





```

```{r Adm_Disch_Country, echo=FALSE, fig.width=11, out.width='100%', warning=FALSE}
# Note that the `echo = FALSE` parameter was added to the code chunk 
# to prevent printing of the R code that generated the plot.

### Plot Admissions, Discharges and Change in Total Patients for the whole Country

mb <- as.numeric(1:9 %o% 10 ^ (1:4))

p <- country %>%
  filter(Region == "All" & Province == "All" & 
           (stringr::str_detect(Category, "(^NEW_|^TOTAL_IN_CHG)")) 
         ) %>% 
  ggplot(aes(x = Date, y = Count, colour = Category)) +
  # scale_y_log10(minor_breaks = log10(mb),
  #               # breaks = scales::trans_breaks("log10", function(x) 10^x),
  #               # labels = scales::trans_format("log10", scales::math_format(10^.x))
  #               ) +
  # annotation_logticks(sides="lr", colour = "white",
  #                     short = unit(1, "npc"),
  #                     mid = unit(1, "npc"),
  #                     long = unit(1, "npc")  
  #                     ) +
  expand_limits(y = 1) +
  geom_hline(yintercept = 0) +
  geom_smooth(se=FALSE) +
  geom_line(linetype='dotted') +
  geom_point(aes(shape = Full_reporting)) +
  ggtitle(paste("Belgium - last updated", format(dt, usetz = TRUE)), 
          subtitle = paste0("Source: ", url)) +
  xlab("Date") +
  ylab("Number of Patients") +
  theme(axis.text.x = element_text(angle = 90),
        # panel.grid.minor.y = element_blank(),
        # panel.grid.major.y = element_blank()
        )
  # print(p)
  
```


### Daily Admissions, Discharges and Changes in Total Number of Patients Hospitalized over Belgium and  by Region
```{r Adm_Disch_by_Reg, echo=FALSE, fig.width=11, out.width='100%', warning=FALSE}
# Note that the `echo = FALSE` parameter was added to the code chunk 
# to prevent printing of the R code that generated the plot.


mb <- as.numeric(1:9 %o% 10 ^ (1:4))

for (r in unique(country$Region)) {
  # cat(paste0(r, "\n"))

  (  country %>%
      filter(Region == r & Province == "All" & 
               (stringr::str_detect(Category, "^(NEW_|TOTAL_IN_CHG)")) 
             ) %>% 
      ggplot(aes(x = Date, y = Count, colour = Category)) +
      expand_limits(y = 1) +
      geom_hline(yintercept = 0) +
      geom_smooth(se=FALSE) +
      geom_line(linetype='dotted') +
      geom_point(aes(shape = Full_reporting)) +
      # facet_wrap(~Region) +
      ggtitle(paste(stringr::str_replace(r, "All", "Belgium"), "- Daily Hospital Admissions and Discharges - last updated", format(dt, usetz = TRUE)),
              subtitle = paste0("Source: ", url)) +
      xlab("Date") +
      ylab("Number of Patients") +
      theme(axis.text.x = element_text(angle = 90),
            ) 
  ) %>% print()  # explicit print() needed for the outputs to be generated within a for loop
  
}
```

### Number of Patients Hospitalized in Belgium and by Region


```{r hosp_sum, echo=FALSE}
hosp_sum <- hosp %>% 
  group_by(REGION, DATE) %>%
  summarise_at(c("NR_REPORTING", "TOTAL_IN", "TOTAL_IN_ICU", "TOTAL_IN_RESP", "TOTAL_IN_ECMO", "NEW_IN", "NEW_OUT")
               ,list(sum)
               )
# kable(hosp_sum)
```

```{r TOTAL_IN, echo=FALSE, fig.width=11, out.width='100%'}
# Note that the `echo = FALSE` parameter was added to the code chunk 
# to prevent printing of the R code that generated the plot.


mb <- as.numeric(1:9 %o% 10 ^ (1:4))

ggplot(hosp_sum, aes(x = DATE, y = TOTAL_IN, colour = REGION)) + 
  scale_y_log10(minor_breaks = log10(mb),
                # breaks = scales::trans_breaks("log10", function(x) 10^x),
                # labels = scales::trans_format("log10", scales::math_format(10^.x))
                ) +
  annotation_logticks(sides="lr", colour = "white",
                      short = unit(1, "npc"),
                      mid = unit(1, "npc"),
                      long = unit(1, "npc")
  ) +
  expand_limits(y = 1) +
  geom_smooth() +
  geom_line(linetype='dotted') +
  geom_point() +
  ggtitle(paste("Belgium - Total Number of Patients Hospitalized - last updated", format(dt, usetz = TRUE)), 
          subtitle = paste0("Source: ", url)) + 
  xlab("Date") +
  ylab("Total Patients Hospitalized") +
  theme(axis.text.x = element_text(angle = 90),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank())
```

```{r TOTAL_IN_ICU, echo=FALSE, fig.width=11, out.width='100%'}
ggplot(hosp_sum, aes(x = DATE, y = TOTAL_IN_ICU, colour = REGION)) + 
  geom_smooth() +
  geom_line(linetype='dotted') +
  geom_point() +
  #scale_y_log10() +
  ggtitle(paste("Belgium - Total Number of Patients in Intensive Care - last updated", format(dt, usetz = TRUE)), 
          subtitle = paste0("Source: ", url)) + 
  xlab("Date") +
  ylab("Total Patients in Intensive Care") +
  theme(axis.text.x = element_text(angle = 90))
```


```{r TOTAL_IN_RESP, echo=FALSE, fig.width=11, out.width='100%'}
ggplot(hosp_sum, aes(x = DATE, y = TOTAL_IN_RESP, colour = REGION)) + 
  geom_smooth() +
  geom_line(linetype='dotted') +
  geom_point() +
  # scale_y_log10() +
  ggtitle(paste("Belgium - Total Number of Patients under Respiratory Support - last updated", format(dt, usetz = TRUE)), 
          subtitle = paste0("Source: ", url)) + 
  xlab("Date") +
  ylab("Total Patients under Respiratory Support") +
  theme(axis.text.x = element_text(angle = 90))
```


```{r TOTAL_IN_ECMO, echo=FALSE, fig.width=11, out.width='100%'}
ggplot(hosp_sum, aes(x = DATE, y = TOTAL_IN_ECMO, colour = REGION)) + 
  geom_smooth(se=FALSE) +
  geom_line(linetype='dotted') +
  geom_point() +
  # scale_y_log10() +
  ggtitle(paste("Belgium - Total Number of Patients on ECMO - last updated", format(dt, usetz = TRUE)), 
          subtitle = paste0("Source: ", url)) + 
  xlab("Date") +
  ylab("Total Patients on ECMO") +
  theme(axis.text.x = element_text(angle = 90))
```