---
title: "Covid-19 Hospitalizations in Belgium"
author: "Jean-Michel Bodart"
date: "`r date()`"
output:
  html_document: default
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
# options(encoding = 'UTF-8')
```

Purpose: provide an overview of the evolution of Covid-19-related hospitalizations in Belgium, by Region and Province. 

```{r get_hosp_be, echo=FALSE, warning=FALSE}
suppressPackageStartupMessages({
library(dplyr)
library (tidyr)
library(httr)
library(lubridate)
library(ggplot2)
library(knitr)
library(stringr)
library(textclean)
#library(purrr)
})
reqdate <- format(seq.Date(as.Date(now()), by=-1, length.out=10), "%Y%m%d")
sources <- data.frame(reqdate, 
                      url = paste0("https://epistat.sciensano.be/Data/", 
                                   reqdate, 
                                   "/COVID19BE_HOSP_",
                                   reqdate,
                                   ".csv"),
                      stringsAsFactors = FALSE)
l <- lapply(sources$url, GET)
sources$status <- vapply(l, function(x){http_status(x)$reason}, "")
wh <- which(sources$status=="OK")
sources$lastmodif <- ''
if (length(wh)>0) {
  sources[wh,]$lastmodif <- vapply(l[wh], function(x){x$headers$`last-modified`}, "")
}
loc <- Sys.getlocale(category = "LC_TIME")
invisible(Sys.setlocale(category = "LC_TIME", locale = "C"))
sources$lastmodifdt <- as.POSIXct(NA)
sources$tz <- ''
sources$tz[wh] <- vapply(sources$lastmodif[wh], function(x){strsplit(x, " ", fixed = TRUE)[[1]][6]}, '')
src <- sources[wh,] %>% 
  mutate(lastmodifdt = as.POSIXct(lastmodif, tz=unique(tz), format="%a, %d %b %Y %H:%M:%S"))
src <- src %>%
  mutate(lastmodifdt = with_tz(lastmodifdt, tz="Europe/Brussels"))
# src <- src %>%
#   mutate(lastmodifdt = with_tz(lastmodifdt, tz="GMT"))
#cat(paste0(format(src$lastmodifdt, usetz = TRUE), "\n"))
sources$lastmodifdt[wh] <- src$lastmodifdt 
sources$tz <- NULL
#cat(paste0(format(src$lastmodifdt, usetz = TRUE), "\n"))
sources <- sources %>%
  mutate(lastmodifdt = with_tz(lastmodifdt, tz="GMT"))
#cat(paste0(format(src$lastmodifdt, usetz = TRUE), "\n"))

invisible(Sys.setlocale(category = "LC_TIME", locale = loc))

url <- sources$url[wh[1]]
dt <- sources$lastmodifdt[wh[1]]

```

## Source Data Sets

-----

### COVID19BE_HOSP (HOSPITALISATIONS BY DATE AND PROVINCE) 

Source: `r url`

Data last modified: `r format(dt, usetz = TRUE)`.

Dataset [description](https://epistat.sciensano.be/COVID19BE_codebook.pdf).

#### Source Variables

Variable      | Definition 
------------- | ---------
DATE          | Date of data collection 
PROVINCE      | Province of reporting hospitals (including Brussels) 
REGION        | Region of reporting hospitals  
NR_REPORTING  | Number of hospitals reporting 
TOTAL_IN      | Total number of hospitalized patients at the moment of reporting (prevalence) 
TOTAL_IN_ICU  | Total number of hospitalized patients in ICU at the moment of reporting (prevalence) 
TOTAL_IN_RESP | Total number of hospitalized patients under respiratory support at the moment of reporting (prevalence) 
TOTAL_IN_ECMO | Total number of hospitalized patients on ECMO (Extracorporeal Membrane Oxygenation) at the moment of reporting (prevalence) 
NEW_IN        | Number of new hospital intakes in the last 24h (incidence) 
NEW_OUT       | Number of new hospital discharges (alive) in the last 24h (incidence) 


#### Newly Derived Variables

Variable | Definition 
-------------- | ---------
Full_reporting | Whether the number of hospitals reporting is equal to the maximal number of hospitals reporting 
TOTAL_IN_CHG   | (Total number of hospitalized patients at the moment of reporting) - (Total number of hospitalized patients the previous day)
NEW_DELTA      | (Number of new hospital intakes) - (new hospital discharges (alive) in the last 24h) 

Source code: [Github](https://github.com/jmbo1190/Covid19)

Summarized data: [csv](https://raw.githubusercontent.com/jmbo1190/Covid19/master/COVID19BE_HOSP_sum.csv) [json](https://raw.githubusercontent.com/jmbo1190/Covid19/master/COVID19BE_HOSP_sum.json)

-----


### Belgium Population by Province

Source: Official data from https://bestat.statbel.fgov.be/bestat/

C01.1: Population par lieu de r&eacute;sidence (province), sexe, position dans le m&eacute;nage (C), &eacute;tat civil et &acirc;ge (B) 

[csv](https://bestat.statbel.fgov.be/bestat/api/views/0f14f470-9a59-4011-ac98-92ebe2df31ff/result/CSV)
[json](https://bestat.statbel.fgov.be/bestat/api/views/0f14f470-9a59-4011-ac98-92ebe2df31ff/result/JSON)

-----

```{r pop_be, echo=FALSE, warning=FALSE}
pop0 <- data.table::fread("https://bestat.statbel.fgov.be/bestat/api/views/0f14f470-9a59-4011-ac98-92ebe2df31ff/result/CSV", 
                         encoding="UTF-8",
                         showProgress = FALSE)
colnames(pop0) <- c("Country", "Region", "Province", "Total", "Household_Position", "Population")

pop0$Region <- replace_non_ascii(pop0$Region)
pop0$Province <- replace_non_ascii(pop0$Province)
prov_en <- c("Bruxelles-Capitale" = "Brussels", 
             "Province d’Anvers" = "Antwerpen", 
             "Anvers" = "Antwerpen", 
             "Limbourg" = "Limburg", 
             "Flandre orientale" = "OostVlaanderen", 
             "Brabant flamand" = "VlaamsBrabant", 
             "Flandre occidentale" = "WestVlaanderen", 
             "Brabant wallon" = "BrabantWallon", 
             "Hainaut" = "Hainaut", 
             "Liege" = "Liege", 
             "Luxembourg" = "Luxembourg", 
             "Namur" = "Namur",
             "All" = "All")

pop1 <- pop0 %>% 
  filter(Household_Position == '') %>% 
  select(-Household_Position, -Total) %>%
  mutate(Country = str_replace(Country, 'Belgique', 'Belgium'),
         Region = str_replace(Region, 'R.+gion de Bruxelles-Capitale', 'Brussels'),
         Region = str_replace(Region, 'R.+gion flamande', 'Flanders'),
         Region = str_replace(Region, 'R.+gion wallonne', 'Wallonia'),
         Region=if_else(Region=='', "All", Region),
         Region=str_squish(Region),
         Province = str_replace(Province, "(R.+gion|Province) (du |de |d')", ''),
         Province=if_else(Province=='', "All", Province),
         Province=prov_en[Province],
         Province=replace_non_ascii(Province)
         ) %>%
  arrange(Population)

pop <- pop1 %>%
  filter(Province != "All") %>%
  group_by(Country, Region) %>%
  summarise(Population = sum(Population)) %>%
  ungroup() %>%
  mutate(Province = "All") %>%
  select(Country, Region, Province, Population) %>%
  rbind(pop1) %>%
  # filter(Province != "All" | Region != "Brussels") %>%
  arrange(Population)
  

```


## Revision History

#### 2020-04-15 - Added Belgium Population Data by Province

Added link to Source Code

Added Figures for:

* Number of Patients Hospitalized per Million Inhabitants by Province
* Number of Patients in Intensive Care per Million Inhabitants by Province
* Number of Patients under Respiratory Support per Million Inhabitants by Province


#### 2020-04-13 - Updated to include Figures by Province

As shown below, the evolution is _not identical_ in all provinces.  


* __As of April 13th:__

  * The total number hospitalized is still on the rise in the province of Namur.
  * The numbers in intensive care are increasing in the provinces of Namur and Oost-Vlaanderen.  
  * The numbers of patients under extracorporeal membrane oxygenation are still on a rising trend in all provinces except Vlaams Brabant, West-Vlaanderen, Li&egrave;ge and Brabant Wallon.

-----

```{r hosp_be, echo=FALSE, warning=FALSE}
# hosp <- data.table::fread("https://epistat.sciensano.be/Data/20200409/COVID19BE_HOSP_20200409.csv")

hosp0 <- data.table::fread(sources$url[wh[1]], showProgress = FALSE) %>%
  mutate(DATE = as.Date(DATE), NEW_DELTA = NEW_IN - NEW_OUT) %>%
  arrange(REGION, PROVINCE, DATE)

# hosp <- hosp0 %>%
#   left_join(
#     select(hosp0, REGION, PROVINCE, DATE, starts_with("TOTAL")) %>%
#       mutate(DATE = DATE + 1) %>%
#       rename_all(
#         list(~ stringr::str_replace_all(., 'TOTAL_', 'PRV_'))
#       ),
#     by = c("REGION", "PROVINCE", "DATE")
#     )  

hosp <- hosp0 %>% 
  group_by(REGION, PROVINCE) %>%
  arrange(DATE, .by_group = TRUE) %>%
  mutate_at(
        vars(starts_with("TOTAL_")), # for all variables starting with "TOTAL_"
        funs(CHG = . - lag(.))       # compute change from previous record
                                     # and store as variable with added suffix "_CHG"
      ) %>%
  ungroup() %>%
  mutate(PROVINCE = str_replace(PROVINCE, "Li.+ge", "Liege")) 
     

hospt <- pivot_longer(data = hosp
                    ,cols = -REGION & -PROVINCE & -DATE & - NR_REPORTING
                    ,names_to = "category"
                    ,values_to = "count") %>%
  mutate(country = "Belgium") %>%
  arrange(country, REGION, PROVINCE, DATE, category) %>%
  select(country, REGION, PROVINCE, DATE, NR_REPORTING, category, count)

province <- hospt %>%
  group_by(country, REGION, PROVINCE, DATE, category) %>%
  summarise(NR_REPORTING = sum(NR_REPORTING), count = sum(count)) %>%
  ungroup() %>%
  select(country, REGION, PROVINCE, DATE, NR_REPORTING, category, count)

region <- hospt %>%
  group_by(country, REGION, DATE, category) %>%
  summarise(NR_REPORTING = sum(NR_REPORTING), count = sum(count)) %>%
  mutate(PROVINCE="All") %>%
  ungroup() %>%
  select(country, REGION, PROVINCE, DATE, NR_REPORTING, category, count) %>%
  rbind(province)

country0 <- hospt %>%
  group_by(country, DATE, category) %>%
  summarise(NR_REPORTING = sum(NR_REPORTING), count = sum(count)) %>%
  ungroup() %>%
  mutate(REGION="All", PROVINCE="All") %>%
  select(country, REGION, PROVINCE, DATE, NR_REPORTING, category, count) %>%
  rbind(region) %>%
  arrange(country, REGION, PROVINCE, DATE, NR_REPORTING, category)

country <- country0  %>%
  group_by(country, REGION, PROVINCE) %>%
  summarize(MAX_REPORTING = max(NR_REPORTING)) %>%
  ungroup() %>%
  full_join(country0, by = c("country", "REGION", "PROVINCE")) %>%
  mutate(full_reporting = (MAX_REPORTING == NR_REPORTING)) %>%
  rename_all(stringr::str_to_title) %>%
  full_join(pop, by = c("Country", "Region", "Province")) %>%
  mutate(Count_per_1M = Count * 1E6 / Population)

write(jsonlite::toJSON(country), 'COVID19BE_HOSP_sum.json')
write.csv(country, file = 'COVID19BE_HOSP_sum.csv')

# country %>% 
#   select(Country, Region, Province, Population) %>%
#   unique()


```

```{r Adm_Disch_Country, echo=FALSE, fig.width=11, out.width='100%', warning=FALSE}
# Note that the `echo = FALSE` parameter was added to the code chunk 
# to prevent printing of the R code that generated the plot.

### Plot Admissions, Discharges and Change in Total Patients for the whole Country

mb <- as.numeric(1:9 %o% 10 ^ (1:4))

p <- country %>%
  filter(Region == "All" & Province == "All" & 
           (stringr::str_detect(Category, "(^NEW_|^TOTAL_IN_CHG)")) 
         ) %>% 
  ggplot(aes(x = Date, y = Count, colour = Category)) +
  # scale_y_log10(minor_breaks = log10(mb),
  #               # breaks = scales::trans_breaks("log10", function(x) 10^x),
  #               # labels = scales::trans_format("log10", scales::math_format(10^.x))
  #               ) +
  # annotation_logticks(sides="lr", colour = "white",
  #                     short = unit(1, "npc"),
  #                     mid = unit(1, "npc"),
  #                     long = unit(1, "npc")  
  #                     ) +
  expand_limits(y = 1) +
  geom_hline(yintercept = 0) +
  geom_smooth(se=FALSE) +
  geom_line(linetype='dotted') +
  geom_point(aes(shape = Full_reporting)) +
  ggtitle(paste("Belgium - last updated", format(dt, usetz = TRUE)), 
          subtitle = paste0("Source: ", url)) +
  xlab("Date") +
  ylab("Number of Patients") +
  theme(axis.text.x = element_text(angle = 90),
        # panel.grid.minor.y = element_blank(),
        # panel.grid.major.y = element_blank()
        )
  # print(p)
  
```


### Daily Admissions, Discharges and Changes in Total Number of Patients Hospitalized over Belgium and by Region
```{r Adm_Disch_by_Reg, echo=FALSE, fig.width=11, out.width='100%', warning=FALSE}
# Note that the `echo = FALSE` parameter was added to the code chunk 
# to prevent printing of the R code that generated the plot.


mb <- as.numeric(1:9 %o% 10 ^ (1:4))

for (r in unique(country$Region)) {
  # cat(paste0(r, "\n"))

  (  country %>%
      filter(Region == r & Province == "All" & 
               (stringr::str_detect(Category, "^(NEW_|TOTAL_IN_CHG)")) 
             ) %>% 
      ggplot(aes(x = Date, y = Count, colour = Category)) +
      expand_limits(y = 1) +
      geom_hline(yintercept = 0) +
      geom_smooth(se=FALSE) +
      geom_line(linetype='dotted') +
      geom_point(aes(shape = Full_reporting)) +
      # facet_wrap(~Region) +
      ggtitle(paste(stringr::str_replace(r, "All", "Belgium"), "- Daily Hospital Admissions and Discharges - last updated", format(dt, usetz = TRUE)),
              subtitle = paste0("Source: ", url)) +
      xlab("Date") +
      ylab("Number of Patients") +
      theme(axis.text.x = element_text(angle = 90),
            ) 
  ) %>% print()  # explicit print() needed for the outputs to be generated within a for loop
  
}
```

------

### Daily Admissions, Discharges and Changes in Total Number of Patients Hospitalized by Province
```{r Adm_Disch_by_Prov, echo=FALSE, fig.width=11, out.width='100%', warning=FALSE}
# Note that the `echo = FALSE` parameter was added to the code chunk 
# to prevent printing of the R code that generated the plot.


mb <- as.numeric(1:9 %o% 10 ^ (1:4))

provinces <- unique(country$Province)
provinces <- provinces[provinces != "All"]

for (P in provinces) {
  # cat(paste0(r, "\n"))

  (  country %>%
      filter(Province == P & 
               (stringr::str_detect(Category, "^(NEW_|TOTAL_IN_CHG)")) 
             ) %>% 
      ggplot(aes(x = Date, y = Count, colour = Category)) +
      expand_limits(y = 1) +
      geom_hline(yintercept = 0) +
      geom_smooth(se=FALSE) +
      geom_line(linetype='dotted') +
      geom_point(aes(shape = Full_reporting)) +
      # facet_wrap(~Province) +
      ggtitle(paste(P, "- Daily Hospital Admissions and Discharges - last updated", format(dt, usetz = TRUE)),
              subtitle = paste0("Source: ", url)) +
      xlab("Date") +
      ylab("Number of Patients") +
      theme(axis.text.x = element_text(angle = 90),
            ) 
  ) %>% print()  # explicit print() needed for the outputs to be generated within a for loop
  
}
```

-----

### Number of Patients Hospitalized in Belgium and by Region


```{r hosp_sum, echo=FALSE}
hosp_sum <- hosp %>% 
  group_by(REGION, DATE) %>%
  summarise_at(c("NR_REPORTING", "TOTAL_IN", "TOTAL_IN_ICU", "TOTAL_IN_RESP", "TOTAL_IN_ECMO", "NEW_IN", "NEW_OUT")
               ,list(sum)
               )
# kable(hosp_sum)
```

```{r TOTAL_IN, echo=FALSE, fig.width=11, out.width='100%'}
# Note that the `echo = FALSE` parameter was added to the code chunk 
# to prevent printing of the R code that generated the plot.


mb <- as.numeric(1:9 %o% 10 ^ (1:4))

ggplot(hosp_sum, aes(x = DATE, y = TOTAL_IN, colour = REGION)) + 
  scale_y_log10(minor_breaks = log10(mb),
                # breaks = scales::trans_breaks("log10", function(x) 10^x),
                # labels = scales::trans_format("log10", scales::math_format(10^.x))
                ) +
  annotation_logticks(sides="lr", colour = "white",
                      short = unit(1, "npc"),
                      mid = unit(1, "npc"),
                      long = unit(1, "npc")
  ) +
  expand_limits(y = 1) +
  geom_smooth() +
  geom_line(linetype='dotted') +
  geom_point() +
  ggtitle(paste("Belgium - Total Number of Patients Hospitalized - last updated", format(dt, usetz = TRUE)), 
          subtitle = paste0("Source: ", url)) + 
  xlab("Date") +
  ylab("Total Patients Hospitalized") +
  theme(axis.text.x = element_text(angle = 90),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank())
```

```{r TOTAL_IN_Reg_per_1M, echo=FALSE, fig.width=11, out.width='100%'}
# Note that the `echo = FALSE` parameter was added to the code chunk 
# to prevent printing of the R code that generated the plot.


mb <- as.numeric(1:9 %o% 10 ^ (1:4))

subset <- country %>% filter(Region != "All" & Province == "All" & Category == "TOTAL_IN")

subset <- subset %>%  mutate(Province=replace_non_ascii(Province))

(ggplot(subset, 
        aes(x = Date, y = Count_per_1M, colour = Region)) + 
  scale_y_log10(minor_breaks = log10(mb),
                # breaks = scales::trans_breaks("log10", function(x) 10^x),
                # labels = scales::trans_format("log10", scales::math_format(10^.x))
                ) +
  annotation_logticks(sides="lr", colour = "white",
                      short = unit(1, "npc"),
                      mid = unit(1, "npc"),
                      long = unit(1, "npc")
  ) +
  expand_limits(y = 1) +
  geom_smooth() +
  geom_line(linetype='dotted') +
  geom_point() +
  ggtitle(paste(r, "- Total Number of Patients Hospitalized per Million Inhabitants by Region - last updated", format(dt, usetz = TRUE)), 
          subtitle = paste0("Source: ", url)) + 
  xlab("Date") +
  ylab("Total Patients Hospitalized per Million") +
  theme(axis.text.x = element_text(angle = 90),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank()
        )
 ) %>% print()
  


```

-----

### Number of Patients Hospitalized in Belgium by Province


```{r hosp_prov_sum, echo=FALSE}
hosp_p_sum <- hosp %>% 
  group_by(REGION, PROVINCE, DATE) %>%
  summarise_at(c("NR_REPORTING", "TOTAL_IN", "TOTAL_IN_ICU", "TOTAL_IN_RESP", "TOTAL_IN_ECMO", "NEW_IN", "NEW_OUT")
               ,list(sum)
               )
# kable(hosp_sum)
```

```{r TOTAL_IN_prov, echo=FALSE, fig.width=11, out.width='100%'}
# Note that the `echo = FALSE` parameter was added to the code chunk 
# to prevent printing of the R code that generated the plot.


mb <- as.numeric(1:9 %o% 10 ^ (1:4))

for (r in c("Flanders", "Wallonia")) {
  
  subset <- hosp_p_sum %>% filter(REGION == r)

(ggplot(subset, 
        aes(x = DATE, y = TOTAL_IN, colour = PROVINCE)) + 
  scale_y_log10(minor_breaks = log10(mb),
                # breaks = scales::trans_breaks("log10", function(x) 10^x),
                # labels = scales::trans_format("log10", scales::math_format(10^.x))
                ) +
  annotation_logticks(sides="lr", colour = "white",
                      short = unit(1, "npc"),
                      mid = unit(1, "npc"),
                      long = unit(1, "npc")
  ) +
  expand_limits(y = 1) +
  geom_smooth() +
  geom_line(linetype='dotted') +
  geom_point() +
  ggtitle(paste(r, "- Total Number of Patients Hospitalized by Province - last updated", format(dt, usetz = TRUE)), 
          subtitle = paste0("Source: ", url)) + 
  xlab("Date") +
  ylab("Total Patients Hospitalized") +
  theme(axis.text.x = element_text(angle = 90),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank()
        )
 ) %>% print()
  
}  
```



```{r TOTAL_IN_prov_per_1M, echo=FALSE, fig.width=11, out.width='100%', warning=FALSE}
# Note that the `echo = FALSE` parameter was added to the code chunk 
# to prevent printing of the R code that generated the plot.


mb <- as.numeric(1:9 %o% 10 ^ (1:4))

for (r in c("Flanders", "Wallonia")) {

  subset <- country %>% filter(Region == r & Category == "TOTAL_IN" & Province != "All")
  
(ggplot(subset, 
        aes(x = Date, y = Count_per_1M, colour = Province)) + 
  scale_y_log10(minor_breaks = log10(mb),
                # breaks = scales::trans_breaks("log10", function(x) 10^x),
                # labels = scales::trans_format("log10", scales::math_format(10^.x))
                ) +
  annotation_logticks(sides="lr", colour = "white",
                      short = unit(1, "npc"),
                      mid = unit(1, "npc"),
                      long = unit(1, "npc")
  ) +
  expand_limits(y = 1) +
  geom_smooth() +
  geom_line(linetype='dotted') +
  geom_point() +
  ggtitle(paste(r, "- Total Number of Patients Hospitalized per Million Inhabitants by Province - last updated", format(dt, usetz = TRUE)), 
          subtitle = paste0("Source: ", url)) + 
  xlab("Date") +
  ylab("Total Patients Hospitalized per Million") +
  theme(axis.text.x = element_text(angle = 90),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank()
        )
 ) %>% print()
  
}  
```

-----

### Number of Patients In Intensive Care in Belgium by Region


```{r TOTAL_IN_ICU_REG, echo=FALSE, fig.width=11, out.width='100%', warning=FALSE}

  (ggplot(country %>% filter(Region != "All" & Province == "All" & Category == "TOTAL_IN_ICU"), 
          aes(x = Date, y = Count, colour = Region)) + 
    geom_smooth() +
    geom_line(linetype='dotted') +
    geom_point() +
    # scale_y_log10() +
    ggtitle(paste("Belgium - Total Number of Patients in Intensive Care by Region - last updated", 
                  format(dt, usetz = TRUE)), 
            subtitle = paste0("Source: ", url)) + 
    xlab("Date") +
    ylab("Total Patients in Intensive Care") +
    theme(axis.text.x = element_text(angle = 90))
  ) %>% print()

```



```{r TOTAL_IN_ICU_REG_per_1M, echo=FALSE, fig.width=11, out.width='100%', warning=FALSE}

  (ggplot(country %>% filter(Region != "All" & Province == "All" & Category == "TOTAL_IN_ICU"), 
          aes(x = Date, y = Count_per_1M, colour = Region)) + 
    geom_smooth() +
    geom_line(linetype='dotted') +
    geom_point() +
    # scale_y_log10() +
    ggtitle(paste("Belgium - Total Number of Patients in Intensive Care per Million Inhabitants by Region - last updated", 
                  format(dt, usetz = TRUE)), 
            subtitle = paste0("Source: ", url)) + 
    xlab("Date") +
    ylab("Total Patients in Intensive Care per Million") +
    theme(axis.text.x = element_text(angle = 90))
  ) %>% print()

```




-----

### Number of Patients In Intensive Care in Belgium by Province


```{r TOTAL_IN_ICU_prov, echo=FALSE, fig.width=11, out.width='100%'}
for (r in c("Flanders", "Wallonia")) {
  
  (ggplot(hosp_p_sum %>% filter(REGION == r), aes(x = DATE, y = TOTAL_IN_ICU, colour = PROVINCE)) + 
    geom_smooth() +
    geom_line(linetype='dotted') +
    geom_point() +
    #scale_y_log10() +
    ggtitle(paste(r, "- Total Number of Patients in Intensive Care by Province - last updated", format(dt, usetz = TRUE)), 
            subtitle = paste0("Source: ", url)) + 
    xlab("Date") +
    ylab("Total Patients in Intensive Care") +
    theme(axis.text.x = element_text(angle = 90))
  ) %>% print()
}
```


```{r TOTAL_IN_ICU_prov_per_1M, echo=FALSE, fig.width=11, out.width='100%', warning=FALSE}
for (r in c("Flanders", "Wallonia")) {
  
  (ggplot(country %>% filter(Region == r & Category == "TOTAL_IN_ICU" & Province != "All"), 
          aes(x = Date, y = Count_per_1M, colour = Province)) + 
    geom_smooth() +
    geom_line(linetype='dotted') +
    geom_point() +
    #scale_y_log10() +
    ggtitle(paste(r, "- Total Number of Patients in Intensive Care per Million Inhabitants by Province - last updated", format(dt, usetz = TRUE)), 
            subtitle = paste0("Source: ", url)) + 
    xlab("Date") +
    ylab("Total Patients in Intensive Care per Million") +
    theme(axis.text.x = element_text(angle = 90))
  ) %>% print()
}
```

-----

### Number of Patients under Respiratory Support in Belgium, by Region and by Province


```{r TOTAL_IN_RESP, echo=FALSE, fig.width=11, out.width='100%'}
ggplot(hosp_sum, aes(x = DATE, y = TOTAL_IN_RESP, colour = REGION)) + 
  geom_smooth() +
  geom_line(linetype='dotted') +
  geom_point() +
  # scale_y_log10() +
  ggtitle(paste("Belgium - Total Number of Patients under Respiratory Support - last updated", format(dt, usetz = TRUE)), 
          subtitle = paste0("Source: ", url)) + 
  xlab("Date") +
  ylab("Total Patients under Respiratory Support") +
  theme(axis.text.x = element_text(angle = 90))
```

```{r TOTAL_IN_RESP_REG_per_1M, echo=FALSE, fig.width=11, out.width='100%', warning=FALSE}

  (ggplot(country %>% filter(Region != "All" & Province == "All" & Category == "TOTAL_IN_RESP"), 
          aes(x = Date, y = Count_per_1M, colour = Region)) + 
    geom_smooth() +
    geom_line(linetype='dotted') +
    geom_point() +
    # scale_y_log10() +
    ggtitle(paste("Belgium - Total Number of Patients under Respiratory Support per Million Inhabitants by Region - last updated", 
                  format(dt, usetz = TRUE)), 
            subtitle = paste0("Source: ", url)) + 
    xlab("Date") +
    ylab("Total Patients under Respiratory Support per Million") +
    theme(axis.text.x = element_text(angle = 90))
  ) %>% print()

```

```{r TOTAL_IN_RESP_prov, echo=FALSE, fig.width=11, out.width='100%'}
for (r in c("Flanders", "Wallonia")) {
  
  (ggplot(hosp_p_sum %>% filter(REGION == r), aes(x = DATE, y = TOTAL_IN_RESP, colour = PROVINCE)) + 
    geom_smooth() +
    geom_line(linetype='dotted') +
    geom_point() +
    # scale_y_log10() +
    ggtitle(paste(r, "- Total Number of Patients under Respiratory Support by Province - last updated", format(dt, usetz = TRUE)), 
            subtitle = paste0("Source: ", url)) + 
    xlab("Date") +
    ylab("Total Patients under Respiratory Support") +
    theme(axis.text.x = element_text(angle = 90))
  ) %>% print()
}
```



```{r TOTAL_IN_RESP_prov_per_1M, echo=FALSE, fig.width=11, out.width='100%', warning=FALSE}
for (r in c("Flanders", "Wallonia")) {
  
  (ggplot(country %>% filter(Region == r & Province != "All" & Category == "TOTAL_IN_RESP"), 
          aes(x = Date, y = Count_per_1M, colour = Province)) + 
    geom_smooth() +
    geom_line(linetype='dotted') +
    geom_point() +
    # scale_y_log10() +
    ggtitle(paste(r, "- Total Number of Patients under Respiratory Support per Million Inhabitants by Province - last updated", 
                  format(dt, usetz = TRUE)), 
            subtitle = paste0("Source: ", url)) + 
    xlab("Date") +
    ylab("Total Patients under Respiratory Support per Million") +
    theme(axis.text.x = element_text(angle = 90))
  ) %>% print()
}
```

-----

### Number of Patients on Extracorporeal Membrane Oxygenation in Belgium, by Region and by Province



```{r TOTAL_IN_ECMO, echo=FALSE, fig.width=11, out.width='100%'}
ggplot(hosp_sum, aes(x = DATE, y = TOTAL_IN_ECMO, colour = REGION)) + 
  geom_smooth(se=FALSE) +
  geom_line(linetype='dotted') +
  geom_point() +
  # scale_y_log10() +
  ggtitle(paste("Belgium - Total Number of Patients on ECMO - last updated", format(dt, usetz = TRUE)), 
          subtitle = paste0("Source: ", url)) + 
  xlab("Date") +
  ylab("Total Patients on ECMO") +
  theme(axis.text.x = element_text(angle = 90))
```

```{r TOTAL_IN_ECMO_prov, echo=FALSE, fig.width=11, out.width='100%'}
for (r in c("Flanders", "Wallonia")) {
  
  (ggplot(hosp_p_sum %>% filter(REGION == r), aes(x = DATE, y = TOTAL_IN_ECMO, colour = PROVINCE)) + 
    geom_smooth(se=FALSE) +
    geom_line(linetype='dotted') +
    geom_point() +
    # scale_y_log10() +
    ggtitle(paste(r, "- Total Number of Patients on ECMO - last updated", format(dt, usetz = TRUE)), 
            subtitle = paste0("Source: ", url)) + 
    xlab("Date") +
    ylab("Total Patients on ECMO") +
    theme(axis.text.x = element_text(angle = 90))
  ) %>% print()
}
```